Solving an Exam Scheduling Problem Using a Genetic Algorithm

### 2009-12-01 trunk/Paper/paper.tex
Paper update, close to finish, few changes from place to place, temporary Lorem ipsum to fill the ending

\section{Approach and Implementation}
  \subsection{Differences From Natural Exam Scheduling Problems}
    The problem we have constructed here is a compromise between the most
    mathematically pure problem with the same sort of properties (which might ignore the
    rooms and related constraints completely) and the problem that must be solved in the
    real world when scheduling thousands of students, which has some important differences
    from this problem.

  \subsection{Tools used}
    The program is implemented in Java, to generate possible schedules and to calculate
    their fitness. A Mathematica program is also written to plot graphs, given the data
    generated by the Java program. Additionally, we use Google code 
    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
    and Google doc to share the work.

@@

    The details of the function we used are essentially arbitrary.

@@

    \subsubsection{General Description}
      The genetic algorithm maintains a population of schedules and constructs subsequent
      generations by adding new random schedules, making copies of high quality schedules 
      in the present generation, applies the mutation operator to schedules, or uses the
      crossover operator on 2 schedules. When performing mutation or crossover, the 
      schedules are selected with a probability proportional to their fitness, allowing 
      the schedules with high quality to be selected more frequently and make more 
      "offspring".

@@

@@

      The following table characterized the problem instances that we used:

      \definecolor{light-gray}{gray}{0.87}
      \rowcolors{1}{light-gray}{white}
      \begin{center}
      \begin{small}
      \begin{tabular}{*{7}{c}}
      \hline 
      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
      \hline \hline
       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
       \textsc{sss}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
      \hline 
	  \end{tabular} 
	  \end{small}
	  \end{center}

	  The large data set was created with parameters inspired from the University of 
	  Toronto's Faculty of Arts and Science.
	  The full input files for these instances are available in our repository in the 
	  directory \emph{Instance Input Files}.

@@

        In the graph below, each data point represents the average of 5 trials. To make 
        the scale clearer, the colour on the surface corresponds to the fitness, where 
        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
        size of the population is fixed, so the number of random individuals used at each 
        point is 60?-(m+c). This corresponds to the distance from the diagonal.

        \begin{figure}[H]
  		  \centering
  		  \includegraphics[width=8cm]{GAParametersData}
  	      \caption{Effectiveness of the GA's proportion parameters}
  	      \label{graph:GAParam1}
		\end{figure}

@@

      \paragraph{Size Parameters}
        Blah blah blaj. We investigated the population size and

@@

  \subsection{Comparison Between the Different Searches}
    [\ldots]
    \begin{figure}[h!]
  	  \centering
  	  \setlength\fboxsep{0pt}
	  \setlength\fboxrule{0.5pt}
	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
      \caption{Search comparison on a large data set}
	\end{figure}

\section{Evaluation}
  \subsection{Success}
    [\ldots]
    Seeing as our fitness function gives a very high penalty for schedules that have direct
    conflicts (either a student with two exams at the same moment, or two different exams
    in a same room at the same moment), if such schedules exists given the input data, 
    our algorithms will tend to very quickly reject them.

\section{Conclusion}
  [\ldots]
  We come up with a conclusion that in this particular problem, genetic algorithm
  is better than random search, and it is better than hill-climbing if the input 
  is a large data set. But it may not be the best solution for this problem, since
  the Mutate Search performs way better.

>>>>

\section{Approach and Implementation}
  \subsection{Differences From Natural Exam Scheduling Problems}
    The problem we have constructed here is a compromise between the most
    mathematically pure problem with the same sort of properties (which might ignore the
    rooms and related constraints completely) and the problem that must be solved in the
    real world when scheduling thousands of students, which has some important differences
    from this problem. For example, we don't deal with constraints such as room capacity
    or exams with varying length.

  \subsection{Tools used}
    Our application was entirely written in Java. We also wrote another Java application 
    to generate problem instance files for testing our algorithms. We used Mathematica 
    to plot all of our graphs.  Additionally, we use Google code's source code
    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}} 
    and Google docs to collaborate on this report.

@@

    The details of the function we used are essentially arbitrary, and we expect 
    similar results would be attained for any other penalty-based fitness function 
    modelling the preferences of students and requirements of rooms.

@@

    \subsubsection{General Description}
      The genetic algorithm maintains a population of schedules and constructs subsequent
      generations by adding new random schedules, making copies of high quality schedules 
      in the present generation, applies the mutation operator to schedules, or uses the
      crossover operator on 2 schedules. When performing mutation or crossover, the 
      schedules are selected with a probability proportional to their fitness, allowing 
      the schedules with high quality to be selected more frequently and make more 
      ``offspring''.

@@

      We will see that this representation is far from ideal, but there is no clear alternative.

@@

      The following table characterized the problem instances that we used 
      (with \textsc{SSS} being an approximation of the search space size):      
      \definecolor{light-gray}{gray}{0.87}
      \rowcolors{1}{light-gray}{white}
      \begin{center}
      \begin{small}
      \begin{tabular}{*{7}{c}}
      \hline 
      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
      \hline \hline
       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
       \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
      \hline 
	  \end{tabular} 
	  \end{small}
	  \end{center}

	  The large data set was created with parameters inspired from the University of 
	  Toronto's Faculty of Arts and Science.

@@

        In the graph below, each data point represents the average of 5 trials. To make 
        the scale clearer, the colour on the surface corresponds to the fitness, where 
        the highest fitnesses are red. The number of copies is fixed at 1, and the proportions 
        of mutations and crossovers are shown on the horizontal axes.

        \begin{figure}[H]
  		  \centering
  		  \includegraphics[width=6cm]{GAParametersData}
  	      \caption{Effectiveness of the GA's proportion parameters}
  	      \label{graph:GAParam1}
		\end{figure}

@@

      \paragraph{Size Parameters}
       [\ldots]  Duis accumsan sagittis velit vel consectetur. Nullam sed ligula nisl. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Morbi quis neque eu justo accumsan auctor. Cras a risus ac est aliquam malesuada quis ac tellus. Aliquam nunc diam, ultrices et cursus ac, ornare mollis tellus. Ut venenatis lacus nec risus eleifend placerat.

@@

		[\ldots]  Praesent eros nulla, ullamcorper non euismod in, pretium et nisi. Nam eu augue ut lacus dignissim gravida. Vivamus euismod convallis lorem, bibendum pellentesque arcu pulvinar a. Phasellus suscipit vehicula nunc non laoreet.

  \subsection{Comparison Between the Different Searches}
    [\ldots] Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum fermentum vehicula tellus, id varius enim cursus nec. In auctor lacus ac diam iaculis id fringilla est tincidunt. Sed tincidunt dictum dolor non dapibus. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus bibendum nisl quis felis accumsan vel hendrerit nulla gravida. Fusce convallis velit eu nisi pellentesque quis sagittis dolor imperdiet. Sed magna velit, convallis in mattis vel, varius vel ante. Nunc euismod aliquam lacus vitae consectetur.

    \begin{figure}[H]
  	  \centering
  	  \setlength\fboxsep{0pt}
	  \setlength\fboxrule{0.5pt}
	  \fbox{\includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}}
      \caption{Comparison of the different searches on a large data set}
	\end{figure}

	[\ldots]Vivamus vehicula metus nibh. Morbi id ante dolor, eget pretium justo. Phasellus sit amet arcu vitae eros malesuada ornare. Proin vulputate volutpat eros, a laoreet ipsum lacinia vel. Vivamus commodo odio non tellus convallis pharetra. Cras varius fringilla tempor. Sed mollis, augue quis euismod pellentesque, turpis mi convallis purus, sed bibendum eros lacus volutpat felis. Quisque velit lectus, dapibus eget tincidunt sed, tincidunt a lectus. Phasellus vel luctus eros. Vestibulum vitae quam mi. Nullam posuere aliquam magna vitae bibendum.

\section{Evaluation}

\section{Conclusion}
  [\ldots]Suspendisse dapibus pretium eros, quis varius nisi tincidunt ac. Duis purus nibh, ullamcorper in ullamcorper eu, pharetra eu nunc. Praesent id ipsum quis est elementum ultricies elementum et odio. Aliquam a mi diam, at posuere lorem. Donec vel nisl lorem. Sed metus nulla, tincidunt sed ornare ac, adipiscing eu augue. Aenean accumsan odio ac sem rutrum feugiat. In hac habitasse platea dictumst. Mauris fringilla dui eget lectus iaculis at venenatis sem tincidunt. Aliquam ullamcorper nibh eget enim consectetur vel bibendum tellus molestie. Etiam mauris lectus, auctor vel ornare eu, accumsan nec enim. Vivamus blandit, est ac vehicula blandit, odio tortor semper lectus, quis luctus eros dolor a elit. Donec consequat tempor purus, eleifend suscipit velit rhoncus vel. Nam sed purus justo, id dapibus urna. Nullam semper mollis tempor.
